### 2024-09-24 paper/paper.qmd
retouch and finish intro and abstract

---
title: "Is Our Community Getting Safer in Recent Years: A Data-Driven Study of Crime Victims in Toronto"
thanks: "Code and data are available at: https://github.com/huayan1998/toronto_crime_analysis-main"
format: pdf
author: Huayan Yu
date: September 24, 2024
editor: visual
date-format: long
abstract: "In this paper, we analyzed crime victim data from Toronto between 2014 and 2023, focusing on various crime subtypes, demographics, and trends. Using visualizations, we examined changes in crime patterns to identify any shifts in public safety conditions. Our findings show a recurring rise in assault cases after 2020 along with a concerning increase in youth and child victims. This study highlights the ongoing challenges in community safety and outlines the need for targeted interventions and policies to address the recent rise in crime."
number-sections: true
bibliography: references.bib
---

```{r}
#| echo: false
#| warning: false
#| message: false

library(bookdown)
library(tidyverse)
library(knitr)
```

# Introduction

Crime and public safety are always important concerns for urban areas, with significant social and economic consequences for local communities. In recent years, technology advancements have enabled researchers to leverage data-driven approaches to better understand and respond to crime patterns. Toronto, as one of Canada's largest and most diverse cities, faces unique challenges in ensuring the safety of its residents. As such, open-source crime data made available by the City of Toronto provides a valuable resource for analyzing crime trends, but certainly more work is needed to translate these patterns into actionable insights for policy and community engagement.

This paper focuses on analyzing crime victim data from Toronto between 2014 and 2023. Specifically, we seek to address the research question: Is the community getting safer in recent years? While numerous studies have explored crime trends, most focus on individual crime types or localized areas. This paper takes a different approach by analyzing various crime subtypes over time, paying particular attention to demographic variables such as sex and age. By visualizing these patterns, we aim to answer the research question quantitatively by identifying potential areas where targeted interventions may be needed.

Our findings show that while most crime types, such as assault, have exhibited a consistent decresing trend in the last decade, they all seem to regain the upward trends in or after 2020, posting new threats and challenges to public safety where youth and child victimization trends are of particular concern. These findings contribute to ongoing public safety discussions in Toronto, emphasizing the need for targeted crime prevention strategies and improved public policy.

In the body of this paper, we will first discuss the data source and our computational environment, along with some basic facts and significances about the data in @sec-data. We then present the analysis results,  focusing on crime subtype trends, gender distribution, and the impact of recent social challenges in @sec-results to address our primary research question for the trend of public safety in recent years in the City of Toronto. Finally in @sec-discuss, we conclude the paper with a summary of key findings and their implications for public safety in Toronto and discuss existing limitations and future directions.

# Data {#sec-data}

## Data Source and Computational Environment

The dataset used in this study, "Police Annual Statistical Report: Victims of Crime," is provided by the City of Toronto through its open data initiative. [1] It reflects crime incident reports, focusing on the experiences of victims from 2019 to 2023. The data is publicly accessible, ensuring transparency in city governance and fostering community engagement in crime monitoring. Ethically, the data adheres to the Open Government License, which mandates privacy protections for individuals involved, promoting responsible use in academic and policy analysis. [2] Statistically, the dataset offers insights into crime trends, aiding in evidence-based policymaking. The data set also provides rich context for statisticians to extract useful insights by analyzing victim components and trends over time.

All data analysis and visualization in this paper were performed using the R programming language (R Core Team 2023) along with the following packages: `tidyverse` (Wickham et al. 2019) which includes the `dplyr` component (Hadley Wickham 2023) for data frame manipulation and `ggplot2` (Wickham 2016) for visualizations, as well as `knitr` (Xie 2023) for table formatting.

## Data Description

The dataset consists of 1244 records from 2014 to 2023, with each column representing critical information as follows:

- `REPORT_YEAR`: The year in which the crime was reported.
- `CATEGORY`: The general classification of the crime, such as "Crimes Against the Person."
- `SUBTYPE`: A more detailed breakdown of the crime, including types like "Assault" or "Robbery."
- `SEX`: Gender of the victim, represented as Male (M), Female (F), or Unknown (U).
- `AGE_GROUP`: The age range of the victim, including "Child", "Adult", "Youth", and "Unknown."
- `AGE_COHORT`: Specific age groupings (e.g., 25 to 34) for more detailed demographic analysis.
- `COUNT_`: The number of occurrences for each record.

In similar studies, other datasets like overall crime reports which usually focus more on offenders could have been used. However, these datasets were not appropriate for our study because this dataset uniquely focuses on victims, providing insights specific to victim demographics and experiences that are important for understanding crime impacts within Toronto local communities.

```{r}
#| echo: false
#| tbl-cap: Number of Records in Each Categorical Variable

# Load the data
data <- read.csv("../data/raw_data/toronto_crime_victims_raw.csv")

# Convert necessary columns to factors
data$CATEGORY <- as.factor(data$CATEGORY)
data$SEX <- as.factor(data$SEX)
data$AGE_COHORT <- as.factor(data$AGE_COHORT)
data$AGE_GROUP <- as.factor(data$AGE_GROUP)
data$SUBTYPE <- as.factor(data$SUBTYPE)
data$ASSAULT_SUBTYPE <- as.factor(data$ASSAULT_SUBTYPE)

# Table 1: data summarization table
# Replace NA values with empty strings for the selected columns
summary_data <- summary(data[,c(3, 4, 6, 8)])
summary_df <- as.data.frame.matrix(summary_data)
summary_df[is.na(summary_df)] <- ""
kable(summary_df, row.names = FALSE)
```

Table 1 above gives an overview of the distribution of key variables, showing that all crimes recorded in the dataset fall under the "Crimes Against the Person" categories. "Assault" is the most common subtype (617), with a fairly even gender distribution (518 females, 570 males). Age cohorts show a diverse range, with "Unknown" being frequent, and many victims in the 25 to 34 age range (165).

The dataset has no missing values in each of its column, and we will not be analyzing the column `ASSAULT_SUBTYPE` because limited information is found in the Open Data Toronto data description page. [1] Hence no further data cleaning is needed other than removing the `ASSAULT_SUBTYPE` column. This comprehensive dataset facilitates focused analysis on victim profiles and crime trends, making it valuable for understanding community safety.

# Results {#sec-results}

## Discovering the victim patterns for different crime subtypes {#sec-results-1}

We first attempt to discover trends in crime over time across different subtypes and identify specific areas that might require targeted interventions. We first consider @fig-one below which focuses on analyzing on crime dynamics in Toronto from 2014 to 2023. Assault has a prominent increasing trend, starting around $15000$ victims in 2014 and rising to over $20000$ by 2023, indicating a substantial escalation in assaults over the past decade except the sudden drop in 2019. On the other hand, crimes categorized as "Other" display a gradual decline after a peak in 2015, stabilizing in recent years, but this subtype is also witnessing a reclimb starting from 2022. Robbery and sexual violations show generally more stable, moderate patterns with minor fluctuations.

This figure reveals that while certain violent crimes, such as assault, have increased, others have remained relatively consistent, highlighting distinct trajectories in various crime types over time. However, all subtypes exhibit upward trends again in or after 2021, adding new uncertainties and concerns about public safety. Considering the fact that assault and sexual violation crimes are usually associated with more physical and psychological injuries, the local community urges some effective policy changes to control escalating crimes.

```{r}
#| label: fig-one
#| fig-cap: Yearly victim trends by subtype from 2014 to 2023
#| echo: false
#| message: false
#| warning: false

# Visualization 1: a line chart of victim subtypes by year (2014 - 2023) 
# Aggregate data by year and subtypes
category_vs_year <- data %>%
  group_by(REPORT_YEAR, SUBTYPE) %>%
  summarize(Total_Count = sum(COUNT_))

# Create line plot
vis1 <- ggplot(category_vs_year, aes(x=REPORT_YEAR, y=Total_Count, color=SUBTYPE)) +
  geom_line(linewidth=1) +
  labs(title="Victim Trends by Subtype Over the Years",
       x="Year", y="Total Victims", color="Crime Subtype") +
  theme_minimal()
vis1
```

As we look closer into the largest crime subtype of assaults in @fig-two, we discover that majority of victims are adults, regardless of gender, with over $70000$ adult victims in both male and female categories. It is worth noting that the victims of unknown age group are at comparative figures as child and youth victims, also posting concerns for the data quality being recorded in the police system. Additionally, while youth and child victims appear to be fewer, they still require further attentions especially after we adjust for the population age distribution.

```{r}
#| label: fig-two
#| fig-cap: Count of victims of assault crimes by sex and age groups
#| echo: false
#| message: false
#| warning: false

# Aggregate data by sex and filtered on assault subtype
assault_victim_by_sex_age <- data %>%
  filter(SUBTYPE == "Assault") %>%
  group_by(SEX, AGE_GROUP) %>%
  summarize(Total_Count = sum(COUNT_))

# Create bar plot
vis2 <- ggplot(assault_victim_by_sex_age, aes(x = SEX, y = Total_Count, fill = AGE_GROUP)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Assault Victim Distribution by Sex and Age Group",
       x = "Sex", y = "Total Victims",
       fill = "Age Group") +
  theme_minimal()
vis2
```

## Discovering the trends of changes in victim counts after 2020 {#sec-results-2}

As we discovered in @sec-results-1, all crime subtypes seem to regain the growth trend starting in or after 2020. This demands a closer examination of the dynamics within the victim data. Table 2 below reveals that the gender distribution of crime victims has remained relatively balanced over the years. This stability in gender distribution suggests a consistent pattern in the way crimes affect male and female populations. More importantly, the two genders both demonstrate a strictly increasing trend in the total number of victims with a growth rate of over $10\%$ from 2021 to 2022. This observation could be evidence for the residual effects of the COVID-19 pandemic and the accompanying economic challenges. The pandemic's disruptions could have contributed to a rise in both opportunity-driven and stress-induced crimes, placing general public mental health condition at another challenge.

In contrast, the unknown category sees a sharp decline, with the total victim count decreasing by 35.23% between 2022 and 2023, indicating improved data collection or identification efforts.

```{r}
#| label: table-two
#| tbl-cap: Proportion and Growth of Victims by Year and Sex
#| echo: false
#| message: false
#| warning: false

proportion_sex_year <- data %>%
  filter(REPORT_YEAR >= 2020) %>%
  group_by(REPORT_YEAR, SEX) %>%
  summarize(Total_Count = sum(COUNT_)) %>%
  group_by(REPORT_YEAR) %>%
  mutate(Proportion = round(Total_Count / sum(Total_Count), 3)) %>%
  arrange(SEX) %>%
  group_by(SEX) %>%
  mutate(Percentage_Change_Total = round((Total_Count - lag(Total_Count)) / lag(Total_Count) * 100, 2))

colnames(proportion_sex_year) <- c("Year", "Sex", "Total", "Proportion",
                                   "Pct Change Total")
proportion_sex_year[is.na(proportion_sex_year)] = 0
kable(proportion_sex_year, row.names=FALSE)
```

As we breakdown the age groups into finer ranges, @fig-three shows that there are consistently a number of outliers in the adult victim category for each year, suggesting that while most adult incidents fall within a consistent range, certain periods saw spikes in crime involving adults, possibly due to heightened vulnerability during specific events. This illustrates the needs for stricter regulations on public events.

Moreover, a slight upward trend in the number of youth and adult victims is observed after 2020, which is likely connected to broader societal impacts, such as increased stress and economic instability following the COVID-19 pandemic. This shift may indicate rising safety concerns faced by the younger populations over time, and calls for a more robust satefy awareness education strategy.

```{r}
#| label: fig-three
#| fig-cap: Victim count distribution by age cohort
#| echo: false
#| message: false
#| warning: false

# Visualization 3: side-by-side boxplot for victims by age groups over year (last 4 years)
vis3 <- ggplot(data %>% filter(REPORT_YEAR >= 2020), 
               aes(x = AGE_GROUP, y = COUNT_, color = AGE_GROUP)) +
  geom_boxplot() +
  labs(title = "Victim Count Distribution by Age Cohort", 
       x = "Age Group", y = "Victim Count") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 60, hjust = 1),
        legend.position = "none") +
  facet_grid(~ REPORT_YEAR)
vis3
```

In Table 3 below, we can more directly observe the significance increase in the number of crime victims across all three age groups. Adult victim counts rise by 19.49%, showing a steady increase. However, the most substantial changes are observed among the youth and child victim groups, with a 75.05% rise in youth victims and a 30.72% increase in child victims. These findings suggest a worrying trend, particularly for younger populations, which might reflect heightened risks for these groups.

```{r}
#| label: table-three
#| tbl-cap: Percentage Changes of Crime Victims over Age Groups
#| echo: false
#| message: false
#| warning: false

victim_change <- data %>%
  filter(REPORT_YEAR %in% c(2020, 2023)) %>%
  group_by(REPORT_YEAR, AGE_GROUP) %>%
  summarize(Total_Victims = sum(COUNT_)) %>%
  spread(REPORT_YEAR, Total_Victims) %>%
  mutate(Pct_Change = round(((`2023` - `2020`) / `2020`) * 100, 2)) %>%
  select(AGE_GROUP, `2020`, `2023`, Pct_Change)
victim_change <- victim_change[c(1, 2, 4),] # remove the unknown category
colnames(victim_change) <- c("Age Group", "2020", "2023", "Percentage Change")
kable(victim_change, row.names=FALSE)
```

# Discussion {#sec-discuss}

## Conclusion

In this study, we start from introducing the open-source dataset that describes crime victims in Toronto in the last ten years, and then proceed with figure and table visualizations to address the research question of whether our community is getting safer in recent years. From our analysis in @sec-results, we discover that assault cases represent the largest proportion of crimes, showing consistent growth over the years. The gender distribution of victims remains balanced, with no significant differences between male and female victimization rates.

However, crime rates across all subtypes began to rise again after 2020 as seen by the increasing numbers of victims across all subtypes, age groups, and genders. The number of youth and child victims also demonstrate upward trends in the new decade with about $75\%$ increase in the number of youth victims which raises particular concerns. Even though recent improvements in data collection methods show positive progress in identifying victims more accurately, public safety is still facing an increasing number of challenges. Based on the data-driven results, we are centainly not confident enough to claim significant progress for the public safety in the Toronto communities.

## Limitations and Next Steps

There are still a few limitations that should be acknowledged in this study. The dataset offers a broad overview without including more specific details on the impacts or contexts of individual crimes, limiting our ability to analyze causes and consequences of crime activities in greater depth. Additionally, the analysis was based on visualization without incorporating advanced statistical models, such as regression analysis or time series forecasting, to better quantify trends or correlations. For instance, cross-factor interactions—like how age and sex may jointly affect crime victimization—remain unexplored.

Future work could address these gaps by obtaining more detailed data on crime contexts, such as socio-economic background or psychological impact assessments, and incorporating insights from professionals in relevant public sectors and social services to provide practical and grounded interpretations. Deploying advanced statistical techniques, such as regression analysis, could further enhance understanding by quantifying correlations and interactions between different factors. This would allow for a deeper and more nuanced exploration of crime trends and help guide policy interventions more effectively.

\newpage

# References

[1]. About Police Annual Statistical Report - Victims of Crimes (https://open.toronto.ca/dataset/police-annual-statistical-report-victims-of-crime/)
[2]. Open Government Licence – Ontario (https://www.ontario.ca/page/open-government-licence-ontario)
[3]. R
[4]. tidyverse
[5]. dplyr
[6]. ggplot2
[7]. knitr